---
title: "Data Manipulation with dplyr"
author: "Farizah Rob"
date: "2022-09-22"
output: github_document
---

Once you've imported data, we're going to need to do some cleaning up. 

```{r}
library(tidyverse)

options(tibble.print_min = 3)

litters_data = read_csv("./data/FAS_litters.csv",
  col_types = "ccddiiii")
litters_data = janitor::clean_names(litters_data)

pups_data = read_csv("./data/FAS_pups.csv",
  col_types = "ciiiii")
pups_data = janitor::clean_names(pups_data)
```

### `select`

You can specify the columns you want to keep by naming all of them.

```{r}
select(litters_data, group, litter_number, gd0_weight, pups_born_alive) 
```

You can also specify column range

```{r}
select(litters_data, group:gd_of_birth) 
```

You can also specify columns you want removed 

```{r}
select(litters_data, -pups_survive) 
select(litters_data, -c(pups_survive, group)) #removing multiple columns together 
```

Rename variables in this process

```{r}
select(litters_data, GROUP = group, LiTtEr_NuMbEr = litter_number) #select columns and change names, first new name=old name
```

If all you want to do is rename something, you can use rename instead of select. This will rename the variables you care about, and keep everything else:

```{r}
rename(litters_data, GROUP = group, LiTtEr_NuMbEr = litter_number)
```
