---
title: "Data Manipulation with dplyr"
author: "Farizah Rob"
date: "2022-09-22"
output: github_document
---

Once you've imported data, we're going to need to do some cleaning up. 

```{r}
library(tidyverse)

options(tibble.print_min = 3)

litters_data = read_csv("./data/FAS_litters.csv",
  col_types = "ccddiiii")
litters_data = janitor::clean_names(litters_data)

pups_data = read_csv("./data/FAS_pups.csv",
  col_types = "ciiiii")
pups_data = janitor::clean_names(pups_data)
```

### `select`

You can specify the columns you want to keep by naming all of them.

```{r}
select(litters_data, group, litter_number, gd0_weight, pups_born_alive) 
```

You can also specify column range

```{r}
select(litters_data, group:gd_of_birth) 
```

You can also specify columns you want removed 

```{r}
select(litters_data, -pups_survive) 
select(litters_data, -c(pups_survive, group)) #removing multiple columns together 
```

Rename variables in this process

```{r}
select(litters_data, GROUP = group, LiTtEr_NuMbEr = litter_number) #select columns and change names, first new name=old name
```

If all you want to do is rename something, you can use rename instead of select. This will rename the variables you care about, and keep everything else:

```{r}
rename(litters_data, GROUP = group, LiTtEr_NuMbEr = litter_number)
```

There are some handy helper functions for select; read about all of them using ?select_helpers. I use starts_with(), ends_with(), and contains() often, especially when there variables are named with suffixes or other standard patterns:

```{r}
select(litters_data, starts_with("gd"))
select(litters_data, ends_with("weight"))
```

I also frequently use is everything(), which is handy for reorganizing columns without discarding anything:

```{r}
select(litters_data, litter_number, pups_survive, everything())
```

`relocate` does a similar thing (and is sort of like rename in that it’s handy but not critical):

```{r}
relocate(litters_data, litter_number, pups_survive)
```

Lastly, like other functions in dplyr, select will export a dataframe even if you only select one column. Mostly this is fine, but sometimes you want the vector stored in the column. To pull a single variable, use `pull`.

Learning Assessment
```{r}
select(pups_data, litter_number, sex, pd_ears)
```


### `filter`

Filter rows according to specific conditions 

Some ways you might filter the litters data are:

* `gd_of_birth == 20`
* `pups_born_alive >= 2`
* `pups_survive != 4`
* `!(pups_survive == 4)`
* `group %in% c("Con7", "Con8")`
* `group == "Con7" & gd_of_birth == 20`

Example

```{r}
filter(litters_data, gd_of_birth==20)
filter(litters_data, group %in% c("Con7", "Con8"))
```


Getting rid of missing values - use the tidyr package instead of filter

```{r}
drop_na(litters_data)
drop_na(litters_data, pups_survive)
```

Don't create too many subsets of the data -> can clutter up the workspace

Learning Assessment
```{r}
filter(pups_data, sex==1)
filter(pups_data, sex==2 & pd_walk<11)
```

### `mutate`


